---
title: ""
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo=FALSE, warning=FALSE, message=FALSE, error=FALSE}
# install.packages("data.table")
# install.packages("RColorBrewer")
# install.packages("cowplot")
 # install.packages("ggplot2")

library(data.table)
library(knitr)
library(RColorBrewer)
library(cowplot)
library(ggplot2)

rm(list=ls())
```

```{r echo=FALSE, error =FALSE, message=FALSE, warning=FALSE}
#1. Importing data 
DT_accuracy <-  fread("PML_accuracy_data_unID.csv")

#2. Adjusting variable types
factor_vars <- c("sex", "pml.category")
DT_accuracy[,factor_vars] <- DT_accuracy[,lapply(.SD, factor), .SDcols=factor_vars]
DT_accuracy[,"age.at.diagnosis"] <- DT_accuracy[,lapply(.SD, function(column){as.integer(gsub(",",".", column))}), .SDcols="age.at.diagnosis"]

#3. Aggregating data for plotting
DT_accuracy_agg_long <- DT_accuracy[`year for diagnose`>=1988,.(.N), by=c("pml.category","year for diagnose")]
DT_accuracy_agg_wide <- dcast(DT_accuracy_agg_long,...~pml.category, value.var="N")[,total.pat.year:=not_pml+pml][,`:=`(not_pml.frac=not_pml/total.pat.year, pml.frac=pml/total.pat.year)]

#4. Plotting 
pml_accuracy_plot <- ggplot(DT_accuracy_agg_wide, aes(x=`year for diagnose`, y=pml.frac))+
  geom_smooth(linetype=2, color="black", fill=brewer.pal(11, "RdBu")[11], alpha=0.5, size=3)+
  geom_point(size=6, color="black", alpha=0.8)+
  xlab("Year (for PML diagnosis)")+
  ylab("Accuracy of PML diagnosis (%)")+
  geom_segment(aes(x=2010, xend=2010, y=0, yend=1), size=2)+
  
  scale_y_continuous(breaks=seq(0,1,0.1), limits=c(0,1), labels = seq(0,100,10))+
  scale_x_continuous(breaks=seq(1988,2014,2), limits=c(1988, 2014))+
  theme(axis.title = element_text(face="bold", size=17))+
  annotate("text", x=1988+6, y=1, label=paste("Average accuracy\n(1988-2010):",toString(round(DT_accuracy_agg_wide[`year for diagnose`<=2010, mean(pml.frac)],2)*100),"%"), size=5, fontface=2)+
  annotate("text", x=1988+25, y=1, label=paste("Average accuracy\n(2011-2013):",toString(round(DT_accuracy_agg_wide[`year for diagnose`>2010, mean(pml.frac)],2)*100),"%"), size=5, fontface=2)

DT_accuracy_agg_wide <- DT_accuracy_agg_wide[,year.group:=sapply(as.list(DT_accuracy_agg_wide[,`year for diagnose`]), function(year){
  if(year<1993){
    return("1988:1992") 
  } else if(year >=1993 & year < 1998){
    return("1993:1997")
  } else if(year >=1998 & year <2003){
    return("1998:2002")
  } else if(year >=2003 & year <2008){
    return("2003:2007")
  } else if (year >=2008 & year <=2013){
    return("2008:2013")
  }
})
][, year.group:=factor(year.group)]

```

###Accuracy of PML Diagnosis 1988:2013
```{r echo=FALSE, error=FALSE, message=FALSE, warning=FALSE, fig.width=12, fig.height=6}
pml_accuracy_plot
```

### Contingency Table of Correct and False PML Diagnosis for Year-Groups 
```{r echo=FALSE, error=FALSE, message=FALSE, warning=FALSE}
pml_contingency <- DT_accuracy_agg_wide[,.(pml=sum(pml), not_pml=sum(not_pml)), by="year.group"]

contingency_table_test <- data.table(parameter="P-value",fisher.test(pml_contingency[,.(pml, not_pml)])$p.value, chisq.test(pml_contingency[,.(pml, not_pml)])$p.value)
names(contingency_table_test) <- c(" ","Fisher's Exact Test", "Pearson's Chi-squared test")

####SECTION OUTPUT
pml_contingency_out <- pml_contingency
pml_contingency_out <- pml_contingency_out[,total.pml:=pml+not_pml]
pml_contingency_out <- rbind(pml_contingency_out,data.table("year.group"="SUM", pml=pml_contingency_out[,sum(pml)], not_pml=pml_contingency_out[,sum(not_pml)], total.pml=pml_contingency_out[,sum(total.pml)]))
pml_contingency_out <- pml_contingency_out[,pml.frac:=round(pml/total.pml,4)]

kable(pml_contingency_out, align=c("l", "c", "c", "c"), col.names = c("Year-Group", "PML", "Not PML", "Total PML", "PML Fraction"))
```

###Statistical Evaluation I: Dependency Between Year-Group and Correct (PML) and False Diagnosis (Not-PML)
```{r echo=FALSE, error=FALSE, message=FALSE, warning=FALSE}
kable(format(contingency_table_test, scientific=T, digits=2), align="c")
```

**Comment:** Table reports p-values for two types of evaluations of dependency in a 2-dimensional contingency table. Null hypothesis is that there is no dependency between the columns (pml/not pml) and the rows (year-groups).

###Statistical Evaluation II: Post-Hoc Analysis of Detected Dependency in Contingency Table
```{r echo=FALSE, error=FALSE, message=FALSE, warning=FALSE}
#1. Post hoc testing fishers and pearsons test of continency table 
comb_list <- list(c(1,2), c(1,3), c(1,4), c(1,5), c(2,3), c(2,4), c(2,5), c(3,4), c(3,5), c(4,5))
#2. Calculating fishers exact test for two year-groups at a time  
post_hoc_table <- data.table(do.call(rbind, lapply(comb_list, function(combination){cbind(t(pml_contingency[combination,1]), format(fisher.test(pml_contingency[combination,2:3])$p.value, nsmall=3, digits = 3))})))
names(post_hoc_table) <- c("year.group.1", "year.group.2", "p.value.unadjusted")
post_hoc_table[, "p.value.unadjusted"] <- as.numeric(post_hoc_table[,p.value.unadjusted])
#3. Evaluating if significant using bonferroni correction 
post_hoc_table <- post_hoc_table[,significant.logical.0.05.level:=ifelse(p.value.unadjusted<=0.05/10,"significant", "not.significant")]

####SECTION OUTPUT
kable(post_hoc_table, align="c", col.names = c("Year.Group.1", "Year.Group.2", "P.value.unadjusted", "Significant.logical"))
```

**Comment:** Two year groups are compared in a 2x2 contingency table at a time using Fisher's exact test. Unadjusted p-values are evaluated using Bonferroni correction at the 5 % level, i.e. test is declared significant if p-value is <=0.05/10 (since we have evaluated 10 combinations). Pearson's chi-sqaured test yielded the same result as Fisher's exact test as was therefore not included. 

###Statistical Evaluation III: Confidence Intervals of Year-Groups
```{r echo=FALSE, error=FALSE, message=FALSE, warning=FALSE}
confidence_out <- DT_accuracy_agg_wide[,.(Mean=mean(pml.frac), SD=sd(pml.frac), N=.N), by="year.group"][,SEMx1.96:=SD/sqrt(N)][,`:=`(CI.lower = Mean-SEMx1.96, CI.upper = Mean+SEMx1.96)][,!c("SD", "N", "SEMx1.96")]

kable(data.table(confidence_out[,1], round(confidence_out[,2:4],2)), align = c("l", rep("c",3)), col.names = c("Year-group", "Mean (PML/Total_PML)", "CI.Lower Bound", "CI.Upper Bound"))
```

**Comment:** Table reports mean (with lower and upper bounds for a 95 % confidence interval) for the fraction of PML/total_PML (i.e. correct diagnosis).
